Blackbox variational inference using \\ the Gaussian process latent variable model

### 2016-02-20 icml2016/paper.tex
changed diag to log diag

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% ICML 2016 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Use the following line _only_ if you're still using LaTeX 2.09.
%\documentstyle[icml2016,epsf,natbib]{article}
% If you rely on Latex2e packages, like most moden people use this:
\documentclass{article}

% use Times
\usepackage{times}
% For figures
\usepackage{graphicx} % more modern
%\usepackage{epsfig} % less modern
\usepackage{subfigure}

@@

% Packages hyperref and algorithmic misbehave sometimes.  We can fix
% this with the following command.
\newcommand{\theHalgorithm}{\arabic{algorithm}}

% Employ the following version of the ``usepackage'' statement for
% submitting the draft version of the paper for review.  This will set
% the note in the first column to ``Under review.  Do not distribute.''
\usepackage{icml2016}

% Employ this version of the ``usepackage'' statement after the paper has
% been accepted, when creating the final version.  This will set the
% note in the first column to ``Proceedings of the...''
%\usepackage[accepted]{icml2016}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Submission and Formatting Instructions for ICML 2016}

@@

\twocolumn[
\icmltitle{Submission and Formatting Instructions for \\
           International Conference on Machine Learning (ICML 2016)}

% It is OKAY to include author information, even for blind
% submissions: the style file will automatically remove it for you
% unless you've provided the [accepted] option to the icml2016
% package.
\icmlauthor{Your Name}{email@yourdomain.edu}
\icmladdress{Your Fantastic Institute,
            314159 Pi St., Palo Alto, CA 94306 USA}
\icmlauthor{Your CoAuthor's Name}{email@coauthordomain.edu}
\icmladdress{Their Fantastic Institute,
            27182 Exp St., Toronto, ON M6H 2T1 CANADA}

% You may provide any keywords that you
% find helpful for describing your paper; these are used to populate
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{boring formatting information, machine learning, ICML}

\vskip 0.3in
]

@@

\section{Introduction}
\label{sec:intro}

@@

\section{Background principles}
\label{sec:background}

\subsection{Variational Inference}
\label{subsec:vi}

@@

\subsection{The Bayesian GP-LVM}
\label{subsec:gplvm}

@@

Denoting the $d^{th}$ column of the data matrix, $Y$, as $\mathbf{y}_d$ the likelihood function for the GP-LVM is
%
\begin{align}
  p(Y|X) &= \prod_{d=1}^D p(\mathbf{y}_d | X),\\
  p(\mathbf{y}_d) &= \mathcal{N}(\mathbf{y}_d|\mathbf{0}, K_{ff}+\beta^{-1}I)
\end{align}
%
where the GPs are modelled as independent across the data dimensions.

@@

\section{GP-LVM-Conditional Mean Field Variational Inference} % (fold)
\label{sec:gplvmcmf}

@@

We begin by applying the standard variational principle to lower bound the marginal log likelihood as follows
%
\begin{align}
  \log p_\theta(Y) &= \mathbb{E}_{q_\phi(Z)}[\log p_\theta(Y|Z) - \log q_\phi(Z)]\notag\\
  &+ \text{KL}[q_\phi(Z)\Vert p_\theta(Z|Y)]\label{eq:var_principle}\\
  &\geq \mathbb{E}_{q_\phi(Z)}[\log p_\theta(Y|Z) - \log q_\phi(Z)] =: \mathcal{L}_1\label{eq:bound1}
\end{align}
 %
 where we denote the set of all fixed point generative (resp. inference) model parameters to be learned by $\theta$ (resp. $\phi$), and we denote the set of all latent variables and parameters subject to inference by $Z$ (though later we will use $\eta$ to ...).
 Integrating \ref{eq:bound1} with respect to the auxiliary variables, noting that the marginal variational distribution over $Z$ can be expressed as $q(Z) = q(Z|F(),U,X)q(F(),U,X)/q(F(),U,X|Z)$, and introducing an auxiliary model $r_\psi(F(),U,X)$, we arrive at the following \emph{auxiliary lower bound}:
 %
 \begin{align}
  \mathcal{L}_1 &\geq \mathbb{E}_{q(Z,F(),U,X)}\big[ \log p(Y|Z)p(Z)r(F(),U,X|Z)\notag\\
   &- \log q(Z|F(),X)q(F(),U,X) \big]\label{eq:lowerlower}\\
    &+ \mathbb{E}_{q(Z)}\big[ \text{KL}[q(F(),U,X|Z)\Vert r(F(),U,X|Z)] \big] =: \mathcal{L}_{aux}.\label{eq:L_aux}
 \end{align}
%
To explain this bound, we find it useful to refer to $q(F(),U,X)$ (resp. $q(F(),U,X|Z)$) as the \emph{variational prior} (resp.) \emph{variational posterior} distributions over the auxiliary inference variables, the latter of which constitutes a posterior belief over the auxiliary variables having observed a (set of) sample(s), $Z$, generated under the measure $q$ endowed by the inference network.

@@

\subsection{Scalable and Amortised Inference}
\label{subsec:scale}

@@

We can rearrange (\ref{eq:L_aux}) as follows:
%
\begin{align}
  \mathcal{L}_{aux} &= \mathbb{E}_{q(Z)}[ \log p(Y|Z) ] - \mathbb{E}_{q(F(),U,X)}[ \text{KL}[q(Z|F(),U,X)\Vert p(Z)] ]\notag\\
  &- \mathbb{E}_{q(Z)}[ \text{KL}[q(F(),U,X)\Vert r(F(),U,X|Z)] ].\label{eq:L_aux_ABC}.
\end{align}
%
As noted by Kingma and Welling, in the case of Gaussian $p(Z)$ and $q(Z|F(),U,X)$, the first K-L divergence can be evaluated analytically and, but the lower lower means we can get a second analytical term... This was also noticed by Tran et al., in their lower lower bound...

Variational compression ... sparse variational GPs ... set $r(F()|U,X,Z):=q(F()|U,X)$ which means that $r(F(),U,X|Z) = q(F()|U,X)r(U,X|Z)$. Full implications of this??? Reduces configuration space... The aptness of this hinges on the veracity of the assumption that ... are sufficient statistics for ... in expectation under ...

@@

\subsection{Design Choices}
\label{subsec:choices}

@@

\section{Related Work}
\label{sec:related}

@@

\section{Experimental Setup}
\label{sec:experiments}

@@

\section{Discussion and Future Work}
\label{sec:discussion}

@@

>>>>

\documentclass[]{article}
\usepackage{proceed2e}
% Set the typeface to Times Roman
\usepackage{times}

% For figures
\usepackage{graphicx} % more modern
%\usepackage{epsfig} % less modern
\usepackage{subfigure}

@@

% Packages hyperref and algorithmic misbehave sometimes.  We can fix
% this with the following command.
\newcommand{\theHalgorithm}{\arabic{algorithm}}
\newcommand{\Ex}{\mathbb{E}}
\newcommand{\KL}{\textrm{KL}}
% I would prefer f or \mathbf{f} rather than F()
\newcommand{\f}{f}

% Just a working title/suggestion
\title{Blackbox variational inference using \\ the Gaussian process latent variable model}

%\author{} % LEAVE BLANK FOR ORIGINAL SUBMISSION.
          % UAI  reviewing is double-blind.

\author{ {\bf Rory Beard} \\
\And
{\bf Chris Lloyd}  \\
\And
{\bf Stephen J. Roberts}   \\
\AND
\textnormal{Machine Learning Research Group} \\
\textnormal{Department of Engineering Science}\\
\textnormal{University of Oxford} \\
\url{{rbeard, clloyd, sjrob}@robots.ox.ac.uk}
}

@@

\maketitle

@@

\section{Introduction}\label{sec:intro}

@@

\section{Background principles}\label{sec:background}

\subsection{Variational Inference}\label{subsec:vi}

@@

\subsection{The Bayesian GP-LVM}\label{subsec:gplvm}

@@

Denoting the $d^{th}$ column of the data matrix, $Y$, as $\mathbf{y}_d$ the likelihood function for the GP-LVM is
%
\begin{align}
  p(Y|X) &= \prod_{d=1}^D p(\mathbf{y}_d | X),\\
  p(\mathbf{y}_d|X) &= \mathcal{N}(\mathbf{y}_d|\mathbf{0}, K_{ff}+\beta^{-1}I)
\end{align}
%
where the GPs are modelled as independent across the data dimensions.

@@

\section{GP-LVM-Conditional Mean Field Variational Inference} \label{sec:gplvmcmf}

@@

We begin by applying the standard variational principle to lower bound the marginal log likelihood as follows
%
\begin{align}
\log p_\theta(Y) &= \Ex_{q_\phi(Z)}[\log p_\theta(Y|Z) - \log q_\phi(Z)]\notag\\
&+ \text{KL}[q_\phi(Z)\Vert p_\theta(Z|Y)]\label{eq:var_principle}\\
&\geq \Ex_{q_\phi(Z)}[\log p_\theta(Y|Z) - \log q_\phi(Z)] \notag \\
&\triangleq \mathcal{L}_1 \label{eq:bound1}
\end{align}
%
where we denote the set of all fixed point generative (resp. inference) model parameters to be learned by $\theta$ (resp. $\phi$), and we denote the set of all latent variables and parameters subject to inference by $Z$ (though later we will use $\eta$ to ...).
Integrating \ref{eq:bound1} with respect to the auxiliary variables, noting that the marginal variational distribution over $Z$ can be expressed as 
\begin{equation}
q(Z) = \frac{q(Z|\f,U,X)q(\f,U,X)}{q(\f,U,X|Z)}
\end{equation}
%
and introducing an auxiliary model $r_\psi(\f,U,X)$, we arrive at the following \emph{auxiliary lower bound}:
%
\begin{align}
\mathcal{L}_1 &\geq \Ex_{q(Z,\f,U,X)}\big[ \log p(Y|Z)p(Z)r(\f,U,X|Z) \notag \\
&- \log q(Z|\f,X)q(\f,U,X) \big] \notag \\
&+ \Ex_{q(Z)}\big[\KL[q(\f,U,X|Z)\Vert r(\f,U,X|Z)]\big] \label{eq:lowerlower} \\
&\triangleq \mathcal{L}_{aux}. \label{eq:L_aux}
\end{align}
%
To explain this bound, we find it useful to refer to $q(\f,U,X)$ (resp. $q(\f,U,X|Z)$) as the \emph{variational prior} (resp.) \emph{variational posterior} distributions over the auxiliary inference variables, the latter of which constitutes a posterior belief over the auxiliary variables having observed a (set of) sample(s), $Z$, generated under the measure $q$ endowed by the inference network.

@@

\subsection{Scalable and Amortised Inference}\label{subsec:scale}

@@

We can rearrange (\ref{eq:L_aux}) as follows:
%
\begin{align}
\mathcal{L}_{aux} &= \Ex_{q(Z)}[ \log p(Y|Z) ] \notag \\
&- \Ex_{q(\f,U,X)}[ \KL[q(Z|\f,U,X)\Vert p(Z)] ]\notag\\
&- \Ex_{q(Z)}[ \KL[q(\f,U,X)\Vert r(\f,U,X|Z)] ].\label{eq:L_aux_ABC}
\end{align}
%
As noted by Kingma and Welling, in the case of Gaussian $p(Z)$ and $q(Z|\f,U,X)$, the first K-L divergence can be evaluated analytically and, but the lower lower means we can get a second analytical term...
This was also noticed by Tran et al., in their lower lower bound...

Variational compression ... sparse variational GPs ... set $r(\f|U,X,Z):=q(\f|U,X)$ which means that $r(\f,U,X|Z) = q(\f|U,X)r(U,X|Z)$.
Full implications of this??? Reduces configuration space...
The aptness of this hinges on the veracity of the assumption that ... are sufficient statistics for ... in expectation under ...

@@

\subsection{Design Choices}\label{subsec:choices}

@@

\section{Related Work}\label{sec:related}

@@

\section{Experimental Setup}\label{sec:experiments}

@@

\section{Discussion and Future Work}\label{sec:discussion}

@@

change!!!!@

\bibliography{paper.bib}{}
\bibliographystyle{plainnat}
